Clique and Constraint Models for Maximum Common (Connected) Subgraph Problems

### 2016-03-30 paper/paper.tex
More progress

% \usepackage{showframe}

\usepackage{microtype}
\usepackage{complexity}                % \P, \NP etc
\usepackage{tikz}                      % For pretty pictures
\usepackage{amsmath}                   % \operatorname
\usepackage{hyperref}                  % clicky links
\usepackage{cleveref}                  % no need to type Figure etc
\usepackage{gnuplot-lua-tikz}          % graphs
\usepackage[ruled,vlined]{algorithm2e} % algorithms (after cleverref!)
\usepackage{placeins}                  % \FloatBarrier
\usepackage{cite}
\usepackage[bottom]{footmisc}

@@

            \draw [very thick] (M1) -- (M2);
            \draw [very thick] (M2) -- (M4);
            \draw [very thick] (M3) -- (M5);
            \draw [very thick] (M4) -- (M6);
            \draw [very thick] (M5) -- (M6);
            \draw [very thick] (M3) -- (M1);
            \draw [very thick] (M6) -- (M7);
            \draw [draw=uofgsandstone!40] (M7) -- (M8);
            \draw [draw=uofgsandstone!40] (M8) -- (M9);
        \end{scope}
\end{tikzpicture}
\caption{A maximum common induced subgraph of the first two graphs has eight vertices. There are
four symmetric solutions, all involving the shaded vertices. However, if we require the common
subgraph to be connected, only seven vertices may be selected---one way to do this is shown in the
third and fourth graphs.}\label{figure:mcsexample}
\end{figure}

@@

A \emph{graph} $G = (V, E)$ consists of a set of vertices $V$ and a set of edges $E \subseteq V
\times V$. In this paper our graphs may be directed, may have loops, and there is a label associated
with each vertex and with each edge. For convenience, we treat undirected graphs as being directed
graphs with symmetric adjacencies, and for unlabelled graphs we treat each vertex as having the same
label, and each edge as having the same label. We write $\operatorname{V}(G)$ for $V$, and
$\operatorname{N}(G, v)$ for the \emph{undirected neighbourhood} of a vertex $v$ (that is, the set
of vertices $w$ such that either $(v, w)$ or $(w, v)$ is an edge).

Given two graphs $G_1$ and $G_2$, we say two pairs of vertices $(v_1, w_1) \in \operatorname{V}(G_1)
\times \operatorname{V}(G_1)$ and $(v_2, w_2) \in \operatorname{V}(G_2) \times
\operatorname{V}(G_2)$ are \emph{compatible} if both $v_1$ and $w_1$ and $v_2$ and $w_2$ are either
non-adjacent or adjacent with the same label, and if the same holds for $(w_1, v_1)$ and $(w_2,
v_2)$.  We say two vertices $v \in \operatorname{V}(G_1)$ and $w \in \operatorname{V}(G_2)$ are
\emph{compatible} if they share the same label, and $(v, v)$ is compatible with $(w, w)$.

?? Get edge vs vertex in early.

@@

One approach to maximum common subgraph problems is via constraint programming
\cite{DBLP:conf/mco/VismaraV08,DBLP:conf/cp/NdiayeS11}. The models are based upon constructing an
injective mapping from the vertices of one graph (typically the smaller) to the other, in a similar
manner to subgraph isomorphism algorithms
\cite{DBLP:journals/ai/Solnon10,DBLP:conf/cp/AudemardLMGP14,DBLP:conf/cp/McCreeshP15}. However, to
allow for a partial mapping, a special $\bot$ value is added to each domain. The objective is then
to minimise the number of variables given the $\bot$ value---for best results, a special propagator
for the injectivity condition is used, which simultaneously enforces that all variables taking
non-$\bot$ are different, and restricts the number of $\bot$ values which may be used based upon the
objective \cite{DBLP:conf/cp/PetitRB01}.  Unfortunately, the addition of $\bot$ also invalidates
many of the invariants used for subgraph isomorphism---we may no longer reason using the degree of
vertices, or the distances between them, for example, and so maximum common subgraph problems can be
much more challenging in practice.

@@

@@

\begin{figure}[p]
    \centering
    \input{gen-graph-unconnected-33}
    \caption{Clique versus constraint programming, 33\% labelled. On the top, the cumulative
    number of instances solved by each approach, in under a certain time. On the bottom, an
    instance-by-instance comparison of the two approaches: the colour at a point indicates the
    number of instances with these runtimes.} \label{figure:unconnected-33}
\end{figure}

\begin{figure}[p]
    \centering
    \input{gen-graph-unconnected-plain}
    \caption{Clique versus constraint programming, undirected, unlabelled. On the top, the cumulative
    number of instances solved by each approach, in under a certain time. On the bottom, an
        instance-by-instance comparison of the two approaches: the colour at a point indicates the
    number of instances with these runtimes.} \label{figure:unconnected-plain}
\end{figure}

@@

The key idea is that whenever we branch (except for the first choice we make), we select only from
variables which correspond to vertices which are adjacent to at least one other variable which has
been instantiated to a value other than $\bot$. If no such variables remain, the subgraph we are building
cannot grow any further, and all remaining variables must be instantiated to $\bot$.

@@

In \cref{algorithm:mccis} we present a clique-inspired algorithm which finds an MCCIS via an
association graph. If the additional
branching restrictions are removed, the core of the algorithm is what Prosser
\cite{DBLP:journals/algorithms/Prosser12} calls the ``MCSa1'' variant of a series of algorithms due
to Tomita \textit{et al.}\
\cite{DBLP:conf/dmtcs/TomitaS03,DBLP:journals/jgo/TomitaK07,DBLP:conf/walcom/TomitaSHTW10}, using a
bitset encoding approach due to San Segundo \textit{et al.}\
\cite{DBLP:journals/cor/SegundoRJ11,DBLP:journals/ol/SegundoMRH13} (and we refer the reader to these
papers for implementation details on how to use bitsets and other data structures to implement the
colouring stage with very low constant factors).

@@

\begin{figure}[p]
    \centering
    \input{gen-graph-connected-undir33}
    \caption{Clique versus constraint programming, connected, undirected, 33\% labelled, up to 100 vertices. On the top,
        the cumulative number of instances solved by each approach, in under a certain time. On the
        bottom, the association graph approach compared to the constraint programming approach with
        both branching and filtering.} \label{figure:connected-undir33}
\end{figure}

\begin{figure}[p]
    \centering
    \input{gen-graph-connected-plain}
    \caption{Clique versus constraint programming, connected, undirected, unlabelled, up to 35 vertices. On the top, the
        cumulative number of instances solved by each approach, in under a certain time. On the
        bottom, the association graph approach compared to the constraint programming approach
    with both branching and filtering.}
\label{figure:connected-plain}
\end{figure}

@@

\FloatBarrier

>>>>

\usepackage{showframe}

\usepackage{microtype}                 % typographic perfection
\usepackage{complexity}                % \P, \NP etc
\usepackage{tikz}                      % For pretty pictures
\usepackage{amsmath}                   % \operatorname
\usepackage{amsfonts}                  % \mathbb
\usepackage{hyperref}                  % clicky links
\usepackage{cleveref}                  % no need to type Figure etc
\usepackage{gnuplot-lua-tikz}          % graphs
\usepackage[ruled,vlined]{algorithm2e} % algorithms (after cleverref!)
\usepackage{cite}                      % automatically sort mass cites
\usepackage[bottom]{footmisc}          % fix footnote + bottom figure on same page

@@

            \draw [very thick] (M1) -- (M2);
            \draw [very thick] (M2) -- (M4);
            \draw [very thick] (M3) -- (M5);
            \draw [very thick] (M4) -- (M6);
            \draw [very thick] (M5) -- (M6);
            \draw [very thick] (M3) -- (M1);
            \draw [very thick] (M6) -- (M7);
            \draw [draw=uofgsandstone!40] (M7) -- (M8);
            \draw [draw=uofgsandstone!40] (M8) -- (M9);
        \end{scope}
\end{tikzpicture}
\caption{A maximum common induced subgraph of the first two graphs has eight vertices, shaded.
However, if we require the common subgraph to be connected, only seven vertices may be
selected---one way to do this is shown in the third and fourth graphs.}\label{figure:mcsexample}
\end{figure}

@@

We will be working with graphs which may be directed, which may have labels associated with
vertices, edges, or both, and which may contain loops. To simplify notation, we will work with a
single definition throughout: we define a \emph{graph} $G = (V, \ell, e)$ to be a set of vertices
$V$, a vertex labelling function $\ell : V \rightarrow \mathbb{N}$, and an edge labelling function
$e : V \times V \rightarrow \mathbb{N} \cup \{ \bot \}$. To represent graphs without vertex (or
edge) labels, we just pick our favourite number and give it to every vertex (or edge). We also use a
distinct label $\bot$ to represent the \emph{lack of} an edge between two vertices.  We write
$\operatorname{V}(G)$ for $V$, and $\operatorname{N}(G, v)$ for the \emph{undirected neighbourhood}
of a vertex $v$ (that is, the set of vertices $w$ such that either $e(v, w) \ne \bot$ or $e(w, v)
\ne \bot$).  A \emph{path} between two vertices $v$ and $w$ is a sequence of $n$ vertices $v = v_1,
v_2, \ldots, v_n = w$ such that for each $m \in \{ 1 \ldots n - 1 \}$, $e(v_m, v_{m + 1}) \ne \bot$.
A graph is \emph{connected} if a path exists between each distinct pair of vertices.

Given two graphs $G_1 = (V_1, \ell_1, e_1)$ and $G_2 = (V_2, \ell_2, e_2)$, we say two pairs of
vertices $(v_1, w_1) \in \operatorname{V}(G_1) \times \operatorname{V}(G_1)$ and $(v_2, w_2) \in
\operatorname{V}(G_2) \times \operatorname{V}(G_2)$ are \emph{compatible}, written $(v_1, w_1)
\simeq (v_2, w_2)$, if both $e_1(v_1, w_1) = e_2(v_2, w_2)$ and $e_1(w_1, v_1) = e_2(w_2, v_2)$. We
say two vertices $v \in \operatorname{V}(G_1)$ and $w \in \operatorname{V}(G_2)$ are
\emph{compatible}, written $v \simeq w$, if $\ell_1(v) = \ell_2(w)$ and $(v, v) \simeq (w, w)$.

An \emph{induced subgraph isomorphism} from a graph $G$ to a graph $H$ is an injective function $i :
\operatorname{V}(G) \rightarrow \operatorname{V}(H)$ such that for every vertex $v \in
\operatorname{V}(G)$, $v \simeq i(v)$, and for every pair of vertices $(v, w) \in
\operatorname{V}(G) \times \operatorname{V}(G)$, $(v, w) \simeq (i(v), i(w))$. A \emph{common
induced subgraph isomorphism} between two graphs $G_1$ and $G_2$ is a graph $G$, together with a
pair of induced subgraph isomorphisms $i_1 : G \rightarrow G_1$ and $i_2 : G \rightarrow G_2$. A
\emph{common induced subgraph} of $G_1$ and $G_2$ is a graph $G$ which has a common subgraph
isomorphism to $G_1$ and $G_2$. A \emph{maximum common (connected) induced subgraph} is a common
(connected) induced subgraph with the largest possible number of vertices.  (The distinction between
a common subgraph and a common subgraph isomorphism is relevant when counting distinct solutions,
which we do not consider in this paper; in \cref{figure:mcsexample} on the left, there is only one
maximum common subgraph, but four different maximum common subgraph isomorphisms are possible.)

There is also a non-induced version of these problems, which is commonly known as the \emph{maximum
common partial subgraph} problem: here, if $e(v, w) = \bot$, then $e_1(i_1(v), i_1(w))$ and
$e_2(i_2(v), i_2(w))$ may take any value, and we must maximise the number of non-$\bot$ vertex pairs
(i.e.\ edges) in $G$ instead (it is not entirely clear what the objective should be in directed
graphs, or when loops are present). We restrict ourselves to the induced version in this paper (??
we could do a journal version, because the clique side is rather involved\ldots).

@@

One approach to maximum common subgraph problems is via constraint programming
\cite{DBLP:conf/mco/VismaraV08,DBLP:conf/cp/NdiayeS11}. The models are based upon constructing an
injective mapping from the vertices of one graph (typically the smaller) to the other, in a similar
manner to subgraph isomorphism algorithms
\cite{DBLP:journals/ai/Solnon10,DBLP:conf/cp/AudemardLMGP14,DBLP:conf/cp/McCreeshP15}. However, to
allow for a partial mapping, a special $\bot$ value is added to each domain (and so we are
constructing a maximum common subgraph isomorphism, where the common subgraph is implicitly the
subgraph of the first graph given by taking the vertices which are not assigned to $\bot$, and the
first subgraph isomorphism is the inclusion map). The objective is then to minimise the number of
variables given the $\bot$ value---for best results, a special propagator for the injectivity
condition is used, which simultaneously enforces that all variables taking non-$\bot$ are different,
and restricts the number of $\bot$ values which may be used based upon the objective
\cite{DBLP:conf/cp/PetitRB01}.  Unfortunately, the addition of $\bot$ also invalidates many of the
invariants used for subgraph isomorphism---we may no longer reason using the degree of vertices, for
example, and so maximum common subgraph problems can be much more challenging in practice.

@@

?? Discuss the partial version briefly, cite the reduction.

@@

\begin{figure}[p]
    \centering
    % \input{gen-graph-unconnected-33}
    \caption{Clique versus constraint programming, 33\% labelled. On the top, the cumulative
    number of instances solved by each approach, in under a certain time. On the bottom, an
    instance-by-instance comparison of the two approaches: the colour at a point indicates the
    number of instances with these runtimes.} \label{figure:unconnected-33}
\end{figure}

\begin{figure}[p]
    \centering
    % \input{gen-graph-unconnected-plain}
    \caption{Clique versus constraint programming, undirected, unlabelled. On the top, the cumulative
    number of instances solved by each approach, in under a certain time. On the bottom, an
        instance-by-instance comparison of the two approaches: the colour at a point indicates the
    number of instances with these runtimes.} \label{figure:unconnected-plain}
\end{figure}

@@

\begin{figure}[tb]
    \newcounter{stepcounter}
    \centering
    \begin{tikzpicture}[scale=0.33]%{{{
        \definecolor{uofgcobalt}{rgb}{0, 0.615686, 0.92549}
        \definecolor{uofgsunshine}{rgb}{1.0, 0.862745, 0.211765}
        \definecolor{uofgthistle}{rgb}{0.584314, 0.070588, 0.447059}
        \definecolor{uofgsandstone}{rgb}{0.321569, 0.278431, 0.231373}
        \begin{scope}
            \node[draw, circle, fill=white, inner sep=0.5pt, font=\normalsize] (Ma) at (0, 4) {\phantom{0}};
            \node[draw, circle, fill=white, inner sep=0.5pt, font=\normalsize] (Mb) at (0, 2) {\phantom{0}};
            \node[draw, circle, fill=white, inner sep=0.5pt, font=\normalsize] (Mc) at (2, 2) {\phantom{0}};
            \node[draw, circle, fill=white, inner sep=0.5pt, font=\normalsize] (Md) at (0, 0) {\phantom{0}};
            \node[draw, circle, fill=white, inner sep=0.5pt, font=\normalsize] (Me) at (2, 0) {\phantom{0}};

            \node [left = 0 of Ma, font=\small] { \vphantom{0}a };
            \node [left = 0 of Mb, font=\small] { \vphantom{0}b };
            \node [right = 0 of Mc, font=\small] { \vphantom{0}c };
            \node [left = 0 of Md, font=\small] { \vphantom{0}d };
            \node [right = 0 of Me, font=\small] { \vphantom{0}e };

            \draw (Ma) -- (Mb);
            \draw (Ma) -- (Mc);
            \draw (Mb) -- (Mc);
            \draw (Mb) -- (Md);
            \draw (Mc) -- (Me);

            \node[draw, circle, fill=white, inner sep=0.5pt, font=\normalsize] (M1) at (5.6, 4) {\phantom{0}};
            \node[draw, circle, fill=white, inner sep=0.5pt, font=\normalsize] (M2) at (7.6, 4) {\phantom{0}};
            \node[draw, circle, fill=white, inner sep=0.5pt, font=\normalsize] (M3) at (5.6, 2) {\phantom{0}};
            \node[draw, circle, fill=white, inner sep=0.5pt, font=\normalsize] (M4) at (7.6, 2) {\phantom{0}};
            \node[draw, circle, fill=white, inner sep=0.5pt, font=\normalsize] (M5) at (5.6, 0) {\phantom{0}};

            \node [left = 0 of M1, font=\small] { \vphantom{0}1 };
            \node [right = 0 of M2, font=\small] { \vphantom{0}2 };
            \node [left = 0 of M3, font=\small] { \vphantom{0}3 };
            \node [right = 0 of M4, font=\small] { \vphantom{0}4 };
            \node [left = 0 of M5, font=\small] { \vphantom{0}5 };

            \draw [color=uofgsandstone!40] (M1) -- (M2);
            \draw [color=uofgsandstone!40] (M1) -- (M3);
            \draw [color=uofgsandstone!40] (M2) -- (M4);
            \draw [color=uofgsandstone!40] (M3) -- (M4);

            \node [anchor=base, yshift=-2.2cm] at ($(Ma)!0.5!(M2)$) {
                \stepcounter{stepcounter}\roman{stepcounter}) Initial problem };
        \end{scope}

        \begin{scope}[xshift=13cm]
            \node[draw, circle, fill=uofgsunshine, inner sep=0.5pt, font=\normalsize] (Ma) at (0, 4) {\phantom{0}};
            \node[draw, circle, fill=white, inner sep=0.5pt, font=\normalsize] (Mb) at (0, 2) {\phantom{0}};
            \node[draw, circle, fill=white, inner sep=0.5pt, font=\normalsize] (Mc) at (2, 2) {\phantom{0}};
            \node[draw, dotted, circle, fill=white, inner sep=0.5pt, font=\normalsize] (Md) at (0, 0) {\phantom{0}};
            \node[draw, dotted, circle, fill=white, inner sep=0.5pt, font=\normalsize] (Me) at (2, 0) {\phantom{0}};

            \node [left = 0 of Ma, font=\small] { \vphantom{0}a };
            \node [left = 0 of Mb, font=\small] { \vphantom{0}b };
            \node [right = 0 of Mc, font=\small] { \vphantom{0}c };
            \node [left = 0 of Md, font=\small] { \vphantom{0}d };
            \node [right = 0 of Me, font=\small] { \vphantom{0}e };

            \draw (Ma) -- (Mb);
            \draw (Ma) -- (Mc);
            \draw (Mb) -- (Mc);
            \draw [dotted] (Mb) -- (Md);
            \draw [dotted] (Mc) -- (Me);

            \node[draw, circle, fill=uofgsunshine, inner sep=0.5pt, font=\normalsize] (M1) at (5.6, 4) {\phantom{0}};
            \node[draw, circle, fill=white, inner sep=0.5pt, font=\normalsize] (M2) at (7.6, 4) {\phantom{0}};
            \node[draw, circle, fill=white, inner sep=0.5pt, font=\normalsize] (M3) at (5.6, 2) {\phantom{0}};
            \node[draw, circle, fill=white, inner sep=0.5pt, font=\normalsize] (M4) at (7.6, 2) {\phantom{0}};
            \node[draw, circle, fill=white, inner sep=0.5pt, font=\normalsize] (M5) at (5.6, 0) {\phantom{0}};

            \node [left = 0 of M1, font=\small] { \vphantom{0}1 };
            \node [right = 0 of M2, font=\small] { \vphantom{0}2 };
            \node [left = 0 of M3, font=\small] { \vphantom{0}3 };
            \node [right = 0 of M4, font=\small] { \vphantom{0}4 };
            \node [left = 0 of M5, font=\small] { \vphantom{0}5 };

            \draw [color=uofgsandstone!40] (M1) -- (M2);
            \draw [color=uofgsandstone!40] (M1) -- (M3);
            \draw [color=uofgsandstone!40] (M2) -- (M4);
            \draw [color=uofgsandstone!40] (M3) -- (M4);

            \node [anchor=base, yshift=-2.2cm] at ($(Ma)!0.5!(M2)$) {
                \stepcounter{stepcounter}\roman{stepcounter}) Try $a \mapsto 1$ };
        \end{scope}

        \begin{scope}[xshift=26cm]
            \node[draw, circle, fill=uofgsunshine, inner sep=0.5pt, font=\normalsize] (Ma) at (0, 4) {\phantom{0}};
            \node[draw, circle, fill=uofgcobalt, inner sep=0.5pt, font=\normalsize] (Mb) at (0, 2) {\phantom{0}};
            \node[draw, circle, fill=white, draw=uofgsandstone!40, inner sep=0.5pt, font=\normalsize] (Mc) at (2, 2) {\phantom{0}};
            \node[draw, circle, fill=white, inner sep=0.5pt, font=\normalsize] (Md) at (0, 0) {\phantom{0}};
            \node[draw, dotted, circle, fill=white, inner sep=0.5pt, font=\normalsize] (Me) at (2, 0) {\phantom{0}};

            \node [left = 0 of Ma, font=\small] { \vphantom{0}a };
            \node [left = 0 of Mb, font=\small] { \vphantom{0}b };
            \node [right = 0 of Mc, font=\small] { \vphantom{0}c };
            \node [left = 0 of Md, font=\small] { \vphantom{0}d };
            \node [right = 0 of Me, font=\small] { \vphantom{0}e };

            \draw [very thick] (Ma) -- (Mb);
            \draw [color=uofgsandstone!40] (Ma) -- (Mc);
            \draw [color=uofgsandstone!40] (Mb) -- (Mc);
            \draw (Mb) -- (Md);
            \draw [color=uofgsandstone!40] (Mc) -- (Me);

            \node[draw, circle, fill=uofgsunshine, inner sep=0.5pt, font=\normalsize] (M1) at (5.6, 4) {\phantom{0}};
            \node[draw, circle, fill=uofgcobalt, inner sep=0.5pt, font=\normalsize] (M2) at (7.6, 4) {\phantom{0}};
            \node[draw, circle, fill=white, inner sep=0.5pt, font=\normalsize] (M3) at (5.6, 2) {\phantom{0}};
            \node[draw, circle, fill=white, inner sep=0.5pt, font=\normalsize] (M4) at (7.6, 2) {\phantom{0}};
            \node[draw, circle, fill=white, inner sep=0.5pt, font=\normalsize] (M5) at (5.6, 0) {\phantom{0}};

            \node [left = 0 of M1, font=\small] { \vphantom{0}1 };
            \node [right = 0 of M2, font=\small] { \vphantom{0}2 };
            \node [left = 0 of M3, font=\small] { \vphantom{0}3 };
            \node [right = 0 of M4, font=\small] { \vphantom{0}4 };
            \node [left = 0 of M5, font=\small] { \vphantom{0}5 };

            \draw [very thick] (M1) -- (M2);
            \draw [color=uofgsandstone!40] (M1) -- (M3);
            \draw [color=uofgsandstone!40] (M2) -- (M4);
            \draw [color=uofgsandstone!40] (M3) -- (M4);

            \node [anchor=base, yshift=-2.2cm] at ($(Ma)!0.5!(M2)$) {
                \stepcounter{stepcounter}\roman{stepcounter}) Try $b \mapsto 2$ };
        \end{scope}

        \begin{scope}[yshift=-9cm]
            \node[draw, circle, fill=uofgsunshine, inner sep=0.5pt, font=\normalsize] (Ma) at (0, 4) {\phantom{0}};
            \node[draw, circle, fill=uofgcobalt, inner sep=0.5pt, font=\normalsize] (Mb) at (0, 2) {\phantom{0}};
            \node[draw, circle, fill=white, draw=uofgsandstone!40, inner sep=0.5pt, font=\normalsize] (Mc) at (2, 2) {\phantom{0}};
            \node[draw, circle, fill=uofgthistle, inner sep=0.5pt, font=\normalsize] (Md) at (0, 0) {\phantom{0}};
            \node[draw, dotted, circle, fill=white, inner sep=0.5pt, font=\normalsize] (Me) at (2, 0) {\phantom{0}};

            \node [left = 0 of Ma, font=\small] { \vphantom{0}a };
            \node [left = 0 of Mb, font=\small] { \vphantom{0}b };
            \node [right = 0 of Mc, font=\small] { \vphantom{0}c };
            \node [left = 0 of Md, font=\small] { \vphantom{0}d };
            \node [right = 0 of Me, font=\small] { \vphantom{0}e };

            \draw [very thick] (Ma) -- (Mb);
            \draw [color=uofgsandstone!40] (Ma) -- (Mc);
            \draw [color=uofgsandstone!40] (Mb) -- (Mc);
            \draw [very thick] (Mb) -- (Md);
            \draw [color=uofgsandstone!40] (Mc) -- (Me);

            \node[draw, circle, fill=uofgsunshine, inner sep=0.5pt, font=\normalsize] (M1) at (5.6, 4) {\phantom{0}};
            \node[draw, circle, fill=uofgcobalt, inner sep=0.5pt, font=\normalsize] (M2) at (7.6, 4) {\phantom{0}};
            \node[draw, circle, fill=white, inner sep=0.5pt, font=\normalsize] (M3) at (5.6, 2) {\phantom{0}};
            \node[draw, circle, fill=uofgthistle, inner sep=0.5pt, font=\normalsize] (M4) at (7.6, 2) {\phantom{0}};
            \node[draw, circle, fill=white, inner sep=0.5pt, font=\normalsize] (M5) at (5.6, 0) {\phantom{0}};

            \node [left = 0 of M1, font=\small] { \vphantom{0}1 };
            \node [right = 0 of M2, font=\small] { \vphantom{0}2 };
            \node [left = 0 of M3, font=\small] { \vphantom{0}3 };
            \node [right = 0 of M4, font=\small] { \vphantom{0}4 };
            \node [left = 0 of M5, font=\small] { \vphantom{0}5 };

            \draw [very thick] (M1) -- (M2);
            \draw [color=uofgsandstone!40] (M1) -- (M3);
            \draw [very thick] (M2) -- (M4);
            \draw [color=uofgsandstone!40] (M3) -- (M4);

            \node [anchor=base, yshift=-2.2cm] at ($(Ma)!0.5!(M2)$) {
                \stepcounter{stepcounter}\roman{stepcounter}) Try $d \mapsto 4$ };
        \end{scope}

        \begin{scope}[xshift=13cm, yshift=-9cm]
            \node[draw, circle, fill=uofgsunshine, inner sep=0.5pt, font=\normalsize] (Ma) at (0, 4) {\phantom{0}};
            \node[draw, circle, fill=uofgcobalt, inner sep=0.5pt, font=\normalsize] (Mb) at (0, 2) {\phantom{0}};
            \node[draw, circle, fill=white, draw=uofgsandstone!40, inner sep=0.5pt, font=\normalsize] (Mc) at (2, 2) {\phantom{0}};
            \node[draw, circle, fill=uofgthistle, inner sep=0.5pt, font=\normalsize] (Md) at (0, 0) {\phantom{0}};
            \node[draw, circle, fill=white, draw=uofgsandstone!40, inner sep=0.5pt, font=\normalsize] (Me) at (2, 0) {\phantom{0}};

            \node [left = 0 of Ma, font=\small] { \vphantom{0}a };
            \node [left = 0 of Mb, font=\small] { \vphantom{0}b };
            \node [right = 0 of Mc, font=\small] { \vphantom{0}c };
            \node [left = 0 of Md, font=\small] { \vphantom{0}d };
            \node [right = 0 of Me, font=\small] { \vphantom{0}e };

            \draw [very thick] (Ma) -- (Mb);
            \draw [color=uofgsandstone!40] (Ma) -- (Mc);
            \draw [color=uofgsandstone!40] (Mb) -- (Mc);
            \draw [very thick] (Mb) -- (Md);
            \draw [color=uofgsandstone!40] (Mc) -- (Me);

            \node[draw, circle, fill=uofgsunshine, inner sep=0.5pt, font=\normalsize] (M1) at (5.6, 4) {\phantom{0}};
            \node[draw, circle, fill=uofgcobalt, inner sep=0.5pt, font=\normalsize] (M2) at (7.6, 4) {\phantom{0}};
            \node[draw, circle, fill=white, inner sep=0.5pt, font=\normalsize] (M3) at (5.6, 2) {\phantom{0}};
            \node[draw, circle, fill=uofgthistle, inner sep=0.5pt, font=\normalsize] (M4) at (7.6, 2) {\phantom{0}};
            \node[draw, circle, fill=white, inner sep=0.5pt, font=\normalsize] (M5) at (5.6, 0) {\phantom{0}};

            \node [left = 0 of M1, font=\small] { \vphantom{0}1 };
            \node [right = 0 of M2, font=\small] { \vphantom{0}2 };
            \node [left = 0 of M3, font=\small] { \vphantom{0}3 };
            \node [right = 0 of M4, font=\small] { \vphantom{0}4 };
            \node [left = 0 of M5, font=\small] { \vphantom{0}5 };

            \draw [very thick] (M1) -- (M2);
            \draw [color=uofgsandstone!40] (M1) -- (M3);
            \draw [very thick] (M2) -- (M4);
            \draw [color=uofgsandstone!40] (M3) -- (M4);

            \node [anchor=base, yshift=-2.2cm] at ($(Ma)!0.5!(M2)$) {
                \stepcounter{stepcounter}\roman{stepcounter}) Forced $e \mapsto \bot$ };
        \end{scope}
    \end{tikzpicture}

    \caption{Suppose we are looking for a connected common subgraph, using the graph on the left
        for variables and the graph on the right (which has an isolated vertex) for values. We
        initially consider $a \mapsto 1$. Our restricted branching rule requires us to select either variable
        $b$ or variable $c$ subsequently, not $d$ or $e$. We try $b \mapsto 2$, which adds $d$
        to the branchable variables, and forces $c \mapsto \bot$. We may now only branch on $d$, and
        we try $d \mapsto 4$. Now the only remaining variable is unbranchable, and so $e = \bot$ is forced, even
        though $5$ remains in its domain.}\label{figure:restricted}
\end{figure}

\cref{figure:restricted} The key idea is that whenever we branch (except for the first choice we
make), we select only from variables which correspond to vertices which are adjacent to at least one
other variable which has been instantiated to a value other than $\bot$. If no such variables
remain, the subgraph we are building cannot grow any further, and all remaining variables must be
instantiated to $\bot$.

@@

In \cref{algorithm:mccis} we present a clique-inspired algorithm which finds a maximum common
connected induced subgraph isomorphism via an
association graph. If the additional
branching restrictions are removed, the core of the algorithm is what Prosser
\cite{DBLP:journals/algorithms/Prosser12} calls the ``MCSa1'' variant of a series of algorithms due
to Tomita \textit{et al.}\
\cite{DBLP:conf/dmtcs/TomitaS03,DBLP:journals/jgo/TomitaK07,DBLP:conf/walcom/TomitaSHTW10}, using a
bitset encoding approach due to San Segundo \textit{et al.}\
\cite{DBLP:journals/cor/SegundoRJ11,DBLP:journals/ol/SegundoMRH13} (and we refer the reader to these
papers for implementation details on how to use bitsets and other data structures to implement the
colouring stage with very low constant factors).

@@

\begin{figure}[p]
    \centering
    % \input{gen-graph-connected-undir33}
    \caption{Clique versus constraint programming, connected, undirected, 33\% labelled, up to 100 vertices. On the top,
        the cumulative number of instances solved by each approach, in under a certain time. On the
        bottom, the association graph approach compared to the constraint programming approach with
        both branching and filtering.} \label{figure:connected-undir33}
\end{figure}

\begin{figure}[p]
    \centering
    % \input{gen-graph-connected-plain}
    \caption{Clique versus constraint programming, connected, undirected, unlabelled, up to 35 vertices. On the top, the
        cumulative number of instances solved by each approach, in under a certain time. On the
        bottom, the association graph approach compared to the constraint programming approach
    with both branching and filtering.}
\label{figure:connected-plain}
\end{figure}

@@
